---
title: "Data for the 2021 Blue Grenadier assessment sent to Patrick Cordue"
author: "Paul Burch"
date: "17 June 2021"
output: 
  html_document:
      toc: TRUE
      toc_float: TRUE
      code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

**This version updates an earlier version with revised age data received from Josh@FAS on 16 June 2021**

This document extracts the Pink Ling age data for Patrick Cordue and compares it
to the data provided to Patrick in 2018.

# Preliminaries

Set default options

```{r}
options(stringsAsFactors = FALSE)
```

Sometimes it's necessary to set a working directory

```{r, eval=FALSE}
# setwd("C:/Work/2020/Projects/Data_services_19in20/sessfdatascripts/Data_ALLspp")
```

Load the packages used in the analysis.  

```{r}
library(SESSFdataproc)
library(SESSFlookups)
library(SESSFmaps)
library(SESSFutils)
```

Get today's date to add to file names of any outputs

```{r}
(date_today <- format(Sys.Date(), "%Y-%m-%d"))
```

Examine the control lookup table to 

```{r}
data("controlLU_2021_06")
controlLU <- controlLU_2021_06
```

# Checks of the Grenadier Ages

Checks and comparisons of the two age data sets

## 2019 Age data

Previous years data brought in from the datapackage 'SESSFdata2019in2020' 
```{r}
data(Age19in20)
```

make a subset for Grenadier
```{r}
dold <- dage[dage$GroupCode %in% "GRE",]
``` 


Examine the data

```{r}
str(dold)
```

Examine the AgeZone field

```{r}
(tbl_zone_dold <- with(dold, table(Year, AgeZone, useNA="always")))
```

Examine the Gear field

```{r}
(tbl_gear_dold <- with(dold, table(Year, Gear, useNA="always")))
```

Examine the LenCode

```{r}
(tbl_len_old <- with(dold, table(Year, LenCode, useNA="always")))
```

## 2021 Age data

Load the CSV created from the spreadsheet  

```{r}
sp_age20in21 <- read.csv("../SESSFdata2020in2021/Data_RAW/Fish_Ageing_Services/Blue Gren Age Update 2021_2021-06-21.csv", header=TRUE)
```

Examine the data

```{r}
str(sp_age20in21)
```

Examine the Zone field by Year

```{r}
(tbl_zone_20in21 <- with(sp_age20in21, table(Year, Zone, useNA="always")))
```

Examine the Gear field by Year

```{r}
(tbl_gear_20in21 <- with(sp_age20in21, table(Year, Gear, useNA="always")))
```

Examine the LenCode

```{r}
(tbl_len_20in21 <- with(sp_age20in21, table(Year, LenCode, useNA="always")))
```

## Comparison of Data 

Compare the number of fish aged by year

```{r}
(age_diffs <- CompareCounts(sp_age15in18$Year, sp_age20in21$Year))
```

Save this as a csv file

```{r}
write.csv(age_diffs, "age_diffs.csv", row.names=FALSE)
```


Compare the zones available in the two years

The 15in18 data

```{r}
with(sp_age15in18, table(AgeZone, useNA="always"))
```

The 20in21 data

```{r}
with(sp_age20in21, table(Zone, useNA="always"))
```

Compare the gear codes

The 15in18 data

```{r}
with(sp_age15in18, table(Gear, useNA="always"))
```

The 20in21 data

```{r}
with(sp_age20in21, table(Gear, useNA="always"))
```

Compare the length codes

The 15in18 data

```{r}
with(sp_age15in18, table(LenCode, useNA="always"))
```

The 20in21 data

```{r}
with(sp_age20in21, table(LenCode, useNA="always"))
```

Write a loop to compare age at length by year

Specify the years to plot

```{r}
(yrs <- sort(unique(sp_age15in18$Year)))
```

**Probably should convert the length types to the reference length for Pink Ling**

```{r}
#opar <- par(mfrow=c(length(yrs),2))

for(i in yrs){
  ## set graphical parameters
  opar <- par(mfrow=c(1,2))
  
  ## create the two length plots for 
  plot(Length ~ Age1, data=sp_age15in18[sp_age15in18$Age1 >0 & sp_age15in18$Length >0 & sp_age15in18$Year %in% i,], main= paste0("Data from 2018 in ", i))
  plot(FloorLength ~ Age1, data=sp_age20in21[sp_age20in21$Age1 >0 & sp_age20in21$FloorLength >0  & sp_age20in21$Year %in% i,], main= paste0("Data from 2021 in ", i))

  ## reset graphical ps
  par(opar)
  
}

#par(opar)

```

# Summary

Fish Ageing Services made a number of corrections to their database in 2019. A 
description of the changes and the queries of these changes is provided in the 
email chain below.

**Email from Paul Burch to Simon@FAS 15 June 2021**

Hi Simon,

I'm preparing the Pink Ling age data for Patrick Cordue who has been doing the Ling assessment the past few times and I just wanted to check that some changes in the data that were probably made in 2019 or 2020 are correct. 

There have been a large number of changes in the numbers of Pink Ling otoliths read between 1993 and 2008 and I just wanted to check that these changes were correct before sending the data to Patrick.

see table to the right

Apologies if we've asked about Ling last year or in 2019 but I could only find emails for other species.

Cheers,
Paul

**email below from Josh@FAS 15 June 2021 in response to a query in changes in Pink Ling age data**

Hi Paul, 

Simon asked me to have a look through the discrepancies in the Ling ageing data between 1993 and 2008. The differences stem from a major database update that occurred a few years ago so that we could include a few extra fields. Including the day of capture became problematic, as dates can vary within batches. 

The vast majority of these discrepancies were updates relating to issues with DOCs, most of which are correct. There were also a few batches (e.g. 256 and 257) where there were some double records that were correctly removed from the DB. 
The only records that were incorrectly updated were from 1993/94, where a number of individuals in a batch had DOCs from different years. I have updated those records. The remaining data post 2008 seem correct, and I have attached the updated data here.

Finally, I have begun going through the same process with blue grens to make sure any discrepancies are correct. My initial look at these suggest that there are a few records that seem problematic. I will get on this tomorrow - along with finalising the age error data also. 

Cheers, 
Josh

**email Paul Burch to Josh@FAS 16 June 2021**

Hi Josh,

Just a couple of queries

There are 26 records without a Year (rows 28785 - 28810)

There's 346 records without a Zone, they appear to come from a number of different trips, some are quite recent but not sure if you can do anything 

There are 71 records that have a length code (LenCode) but don't have a length, again not sure you can do anything.

Thanks again for looking at these

Cheers,
Paul


**There was a followup email from Josh@FAS with corrected data on 16 June 2021**

Hi Paul, 

No worries! 

- The 26 records unfortunately didn't have any DOC data and couldn't be matched up with the other SIDaC data to get a DOC. We do think that they are 2019 - but can't be sure of that (and completely unsure of day/month).

- Unfortunately, most of these did not have any zone data. We did have the data for some of them 
- which I've updated. There are now 210 records with no zone data.

- Missing lengths - there isn't much I can do about most of these unfortunately. I've filled in any blanks that I could find. Some of these, the length data just wasn't there, so the LenCat shouldn't have been there either. Others were part of a batch where length data were there for all individuals except one or two - so i assume it was a drag down issue. 

I've attached the updated data here. 

Let me know if there are any other issues. 
Josh


# Session information

Print the session information

```{r}
sessionInfo()
```

